---
title: "Alpha_diversity"
author: "Wenxuan Dong"
date: '2022-06-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load package, include=FALSE}
library(tidyverse)
library(qiime2R)
library(ggpubr)
library(ggsci)
library(lme4)
library(afex)
my_colors <- c("#DC0000FF","#DC0000F7","#DC0000EF","#DC0000E8","#DC0000E0","#DC0000D8",
              "#DC0000D1","#DC0000C9","#DC0000C1","#DC0000BA","#DC0000B2","#DC0000AA",
              "#DC0000A3","#DC00009B","#DC000093","#DC00008C","#DC000084","#DC00007C",
              "#DC000075","#DC00006D","#DC000066","#DC00005E","#DC000056","#DC00004F",
              "#DC000047","#DC00003F","#DC000038","#DC000030","#DC000028","#DC000021",
              "#DC000019","#3C548819",
              "#3C548821","#3C548828","#3C548830","#3C548838","#3C54883F","#3C548847",
              "#3C54884F","#3C548856","#3C54885E","#3C548866","#3C54886D","#3C548875",
              "#3C54887C","#3C548884","#3C54888C","#3C548893","#3C54889B","#3C5488A3",
              "#3C5488AA","#3C5488B2","#3C5488BA","#3C5488C1","#3C5488C9","#3C5488D1",
              "#3C5488D8","#3C5488E0","#3C5488E8","#3C5488EF","#3C5488F7","#3C5488FF")
```

```{r read tables, include=FALSE}
meta <- read.table("input/metadata.con.rare.txt", sep = '\t')
meta$age_day <- as.numeric(meta$age_day)
meta$age_day_fac <- factor(meta$age_day, levels = c("0","1","2","3","4","5","6","7","10",
                                                    "11","12","13","14","15","16","17","18", "20","21","22","23","24","25","27","28","29","30","32","33","35","39","41","42","49", 
"50","51","55","56","61","63","70","76","77","84","90","91","97","99","104","112","116",
                              "118","130","132","138","140","146","154","155","159","174","183"))
evenness = read_qza("input/core-metrics-results/evenness_vector.qza")
evenness <- evenness$data %>% rownames_to_column("SampleID")
observed_features = read_qza("input/core-metrics-results/observed_features_vector.qza")
observed_features<-observed_features$data %>% rownames_to_column("SampleID")
observed_features$observed_features <- as.numeric(observed_features$observed_features)
shannon = read_qza("input/core-metrics-results/shannon_vector.qza")
shannon<-shannon$data %>% rownames_to_column("SampleID")
faith_pd = read_qza("input/core-metrics-results/faith_pd_vector.qza")
faith_pd<-faith_pd$data %>% rownames_to_column("SampleID")
```

```{r merge tables, include=FALSE}
alpha_diversity = merge(x=faith_pd, y=evenness, by.x = "SampleID", by.y = "SampleID")
alpha_diversity = merge(alpha_diversity, observed_features, by.x = "SampleID", by.y = "SampleID")
alpha_diversity = merge(alpha_diversity, shannon, by.x = "SampleID", by.y = "SampleID")
meta = merge(meta, alpha_diversity, by.x = "sample.id", by.y = "SampleID")
row.names(meta) = meta$SampleID
```

# Check normal distribution

```{r hist & qqnorm, fig.show="hide"}
hist(meta$shannon, main="Shannon diversity", xlab="", breaks=10)
hist(meta$faith_pd, main="Faith phylogenetic diversity", xlab="", breaks=10)
hist(meta$pielou_e, main="Evenness", xlab="", breaks=10)
hist(as.numeric(meta$observed_features), main="Observed Features", xlab="", breaks=10)
ggqqplot(meta$shannon, title = "Shannon")
ggqqplot(meta$faith_pd, title = "Faith PD")
ggqqplot(meta$pielou_e, title = "Evenness")
ggqqplot(meta$observed_features, title = "Observed Features")
```

```{r shapiro test}
shapiro.test(meta$shannon)
shapiro.test(meta$pielou_e)
shapiro.test(meta$faith_pd)
shapiro.test(meta$observed_features)
```

# Alpha diversity analysis over time points

## GLM model
We use a General Linear Model for our **non-normally distributed** metrics. We change the distribution from Gaussian (normal) distribution to **Quasipoisson (log) distribution** that fits the data better.

### Faith
```{r fit models1, fig.show='hold'}
# Gaussian (normal) distribution
gaussian.faith.time = glm(faith_pd ~ age_day + paper, data=meta, family="gaussian")
plot(gaussian.faith.time, which=c(1,2))

# Quasipoisson (log) distribution
qp.faith.time = glm(faith_pd ~ age_day + paper, data=meta, family="quasipoisson")
plot(qp.faith.time, which=c(1,2))
```

post-hoc test
```{r post hoc1}
pair_faith <- pairwise.wilcox.test(meta$faith_pd, meta$age_day, p.adjust.method="BH")
# write.table(pair_faith$p.value, 'output/alpha_faith_pairwise.txt', quote = F,
#             sep = '\t', row.names = T)
```

statistical results 
```{r statistic1, warning=FALSE}
# print the coefficients (age_day) of GLM model
coef(summary(qp.faith.time))['age_day', ]
# linear correlation 
cor.faith <- cor.test(meta$age_day, meta$faith_pd, method = "spearman")
cor.faith
lm.faith <- lm(age_day~faith_pd, meta)
```

box plots shows the Faith PD change over time
```{r ggplot1}

ggplot(data = meta, aes(x=age_day_fac, y=faith_pd, # faith_pd,pielou_evenness,shannon_entropy,observed_features
           fill=age_day_fac), color = "black") +
  geom_boxplot(notch = FALSE,
               #outlier.colour = "gray",
               outlier.shape = 21) +
  stat_boxplot(geom = "errorbar",aes(ymin=..ymax..),
               width=0.5,color="black") +
  stat_boxplot(geom = "errorbar",aes(ymax=..ymin..),
               width=0.5,color="black") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(legend.title = element_blank(),
        legend.position = "none") + 
  scale_fill_manual(values = my_colors) +
  labs(y="Faith Phylogenetic Diversity", x = "Days postnatal",
       title = paste0("Spearman's rho = ", round(cor.faith$estimate, digits = 2), ", ", 
                                                "P = ", signif(cor.faith$p.value, digits = 3), ", ",
                                                "R squared = ", round(summary(lm.faith)$r.squared, digits = 3))) 
# ggsave("output/alpha_age_faith.pdf", height = 4, width = 10)
```

### Pielou's Evenness
```{r fit models2, fig.show='hold'}
# Gaussian (normal) distribution
gaussian.even.time = glm(pielou_evenness ~ age_day + paper, data=meta, family="gaussian")
plot(gaussian.even.time, which=c(1,2))

# Quasipoisson (log) distribution
qp.even.time = glm(pielou_evenness ~ age_day + paper, data=meta, family="quasipoisson")
plot(qp.even.time, which=c(1,2))
```

post-hoc test
```{r post hoc2}
pair_even <- pairwise.wilcox.test(meta$pielou_evenness, meta$age_day, p.adjust.method="BH")
write.table(pair_even$p.value, 'output/alpha_even_pairwise.txt', quote = F,
            sep = '\t', row.names = T)
```

statistical results 
```{r statistic2, warning=FALSE}
# print the coefficients (age_day) of GLM model
coef(summary(qp.even.time))['age_day', ]
# linear correlation 
cor.even <- cor.test(meta$age_day, meta$pielou_evenness, method = "spearman")
cor.even
lm.even <- lm(age_day~pielou_evenness, meta)
```

box plots shows the evenness change over time
```{r ggplot2}

ggplot(data = meta, aes(x=age_day_fac, y=pielou_evenness, # faith_pd,pielou_evenness,shannon_entropy,observed_features
           fill=age_day_fac), color = "black") +
  geom_boxplot(notch = FALSE,
               #outlier.colour = "gray",
               outlier.shape = 21) +
  stat_boxplot(geom = "errorbar",aes(ymin=..ymax..),
               width=0.5,color="black") +
  stat_boxplot(geom = "errorbar",aes(ymax=..ymin..),
               width=0.5,color="black") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(legend.title = element_blank(),
        legend.position = "none") + 
  scale_fill_manual(values = my_colors) +
  labs(y="Pielou's Evenness", x = "Days postnatal",
       title = paste0("Spearman's rho = ", round(cor.even$estimate, digits = 2), ", ", 
                                                "P = ", signif(cor.even$p.value, digits = 3), ", ",
                                                "R squared = ", round(summary(lm.even)$r.squared, digits = 3))) 
# ggsave("output/alpha_age_even.pdf", height = 4, width = 10)
```

### Shannon
```{r fit models3, fig.show='hold'}
# Gaussian (normal) distribution
gaussian.shannon.time = glm(shannon_entropy ~ age_day + paper, data=meta, family="gaussian")
plot(gaussian.shannon.time, which=c(1,2))

# Quasipoisson (log) distribution
qp.shannon.time = glm(shannon_entropy ~ age_day + paper, data=meta, family="quasipoisson")
plot(qp.shannon.time, which=c(1,2))
```

post-hoc test
```{r post hoc3}
pair_shannon <- pairwise.wilcox.test(meta$shannon_entropy, meta$age_day, p.adjust.method="BH")
write.table(pair_shannon$p.value, 'output/alpha_shannon_pairwise.txt', quote = F,
            sep = '\t', row.names = T)
```

statistical results 
```{r statistic3, warning=FALSE}
# print the coefficients (age_day) of GLM model
coef(summary(qp.shannon.time))['age_day', ]
# linear correlation 
cor.shannon <- cor.test(meta$age_day, meta$shannon_entropy, method = "spearman")
cor.shannon
lm.shannon <- lm(age_day~shannon_entropy, meta)
```

box plots shows the Shannon change over time
```{r ggplot3}

ggplot(data = meta, aes(x=age_day_fac, y=shannon_entropy, # faith_pd,pielou_evenness,shannon_entropy,observed_features
           fill=age_day_fac), color = "black") +
  geom_boxplot(notch = FALSE,
               #outlier.colour = "gray",
               outlier.shape = 21) +
  stat_boxplot(geom = "errorbar",aes(ymin=..ymax..),
               width=0.5,color="black") +
  stat_boxplot(geom = "errorbar",aes(ymax=..ymin..),
               width=0.5,color="black") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(legend.title = element_blank(),
        legend.position = "none") + 
  scale_fill_manual(values = my_colors) +
  labs(y="Shannon Diversity", x = "Days postnatal",
       title = paste0("Spearman's rho = ", round(cor.shannon$estimate, digits = 2), ", ", 
                                                "P = ", signif(cor.shannon$p.value, digits = 3), ", ",
                                                "R squared = ", round(summary(lm.shannon)$r.squared, digits = 3))) 
# ggsave("output/alpha_age_shannon.pdf", height = 4, width = 10)
```

### Number of observed features
```{r fit models4,fig.show='hold'}
# Gaussian (normal) distribution
gaussian.observed.time = glm(observed_features ~ age_day + paper, data=meta, family="gaussian")
plot(gaussian.observed.time, which=c(1,2))

# Quasipoisson (log) distribution
qp.observed.time = glm(observed_features ~ age_day + paper, data=meta, family="quasipoisson")
plot(qp.observed.time, which=c(1,2))
```

post-hoc test
```{r post hoc4, warning=FALSE}
pair_observed <- pairwise.wilcox.test(meta$observed_features, meta$age_day, p.adjust.method="BH")
write.table(pair_observed$p.value, 'output/alpha_observed_pairwise.txt', quote = F,
            sep = '\t', row.names = T)
```

statistical results 
```{r statistic4, warning=FALSE}
# print the coefficients (age_day) of GLM model
coef(summary(qp.observed.time))['age_day', ]
# linear correlation 
cor.observed <- cor.test(meta$age_day, meta$observed_features, method = "spearman")
cor.observed
lm.observed <- lm(age_day~observed_features, meta)
```

box plots shows the observed change over time
```{r ggplot4}

ggplot(data = meta, aes(x=age_day_fac, y=observed_features, # faith_pd,pielou_evenness,shannon_entropy,observed_features
           fill=age_day_fac), color = "black") +
  geom_boxplot(notch = FALSE,
               #outlier.colour = "gray",
               outlier.shape = 21) +
  stat_boxplot(geom = "errorbar",aes(ymin=..ymax..),
               width=0.5,color="black") +
  stat_boxplot(geom = "errorbar",aes(ymax=..ymin..),
               width=0.5,color="black") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(legend.title = element_blank(),
        legend.position = "none") + 
  scale_fill_manual(values = my_colors) +
  labs(y="Number of Observed Features", x = "Days postnatal",
       title = paste0("Spearman's rho = ", round(cor.observed$estimate, digits = 2), ", ", 
                                                "P = ", signif(cor.observed$p.value, digits = 3), ", ",
                                                "R squared = ", round(summary(lm.observed)$r.squared, digits = 3))) 
# ggsave("output/alpha_age_observed.pdf", height = 4, width = 10)
```

# Alpha diversity analysis with treatments

## Antibiotics
```{r load data, include=FALSE}
meta <- read.table("input/sample-metadata.tsv", sep = '\t', header = TRUE )
meta = merge(meta, alpha_diversity, by.x = "sample.id", by.y = "SampleID")
row.names(meta) = meta$SampleID
ab_data <- filter(meta, study %in% c('amb21','fi19','fvs20'),
                  treatment %in% c('antibiotic','control'))
```

### Linear Mixed-effects Model
Linear Mixed-effects Model with pig age and treatment as fixed effects, paper and pig id as random effects.
```{r fit modell, include=FALSE}

mixed.time.ab <- lmer(observed_features ~ age_day+treatment+age_day*treatment + (1|paper) + (1|pig_id), 
                   data = ab_data)

anova(mixed.time.ab)
summary(mixed.time.ab)
coef(summary(mixed.time.ab))
confint(mixed.time.ab)

a <- data.frame(coef(summary(mixed.time.ab)))
write.table(a, "output/alpha_ab_observe_mixed.txt", quote = F, sep = "\t",
            row.names = T)
```

post hoc test
```{r post hocc, include=FALSE}
ab_days <- unique(ab_data$age_day)
for (i in ab_days){
  print(i)
  data <- filter(ab_data, age_day == i)
  print(wilcox.test(faith_pd~treatment, data=data))
}

 for (i in ab_days){
  print(i)
  data <- filter(ab_data, age_day == i)
  print(wilcox.test(pielou_evenness~treatment, data=data))
}

for (i in ab_days){
  print(i)
  data <- filter(ab_data, age_day == i)
  print(wilcox.test(shannon_entropy~treatment, data=data))
}

for (i in ab_days){
  print(i)
  data <- filter(ab_data, age_day == i)
  print(wilcox.test(observed_features~treatment, data=data))
}
```

visualize by boxplot
```{r, include=FALSE}
ggboxplot(filter(meta, study %in% c('amb21','fi19','fvs20'),
                 treatment %in% c('antibiotic','control'),
                 age_day != 0), 
          x='age_day', y='faith_pd', 
          color = 'treatment', add='jitter')+ 
          #add='jitter', ylim=c(0,400)) +
  scale_color_manual(breaks = c('control', 'antibiotic'),
                     values = c('#f44336','#2986cc'))+
  labs(y='Faith phylogenetic diversity', x='Days postnatal', color=NULL) +
  theme_minimal() +
  theme(legend.position = c(.9,.2),
        axis.text = element_text(size = 12),
        axis.title.y = element_text(size=12))

```






