---
title: "Metadata"
author: "Wenxuan Dong"
date: '2022-06-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(ggplot2)
library(ggmap)
library(sp)
library(maptools)
library(maps)
library(ggsci)
library(tidyverse)
library(ggpubr)
```

# Map shows the data distribution

To avoid overlap as much as possible, we slightly modified the coordinate of some studies (from Latitude&Longitude to Latitude_new&Longitude_new). But they are still very close to their original location. 

```{r map}
geodata <- read.table("input/metamap.txt", header = TRUE, sep = "\t")
mapworld <- borders("world", colour = "white", fill = "grey90")
ggplot(geodata) + 
  mapworld + 
  ylim(-60, 90) + # remove south pole
  geom_point(aes(
    x = Longitude_new,
    y = Latitude_new,
    color = Study), 
    alpha = 1, size = 2) +
  labs(x="Longitude", y="Latitude") +
  coord_cartesian() +
  scale_color_simpsons()+
  labs(x=NULL,y=NULL)+
  theme_minimal() +
  theme(legend.position = c(.12,.5),
        legend.background = element_blank(),
        legend.key.size = unit(0.5, "cm"),
        axis.text.x=element_blank(),
        axis.text.y=element_blank())
#ggsave("output/metamap.pdf", width = 9, height = 5)
```

# Pie chart shows the data distribution in different catogories

### Country, Hypervariable region, Growth stage, Squencer platform
```{r pie}
meta <- read.table("input/sample-metadata.tsv", header = TRUE, sep = "\t")

a <- ggplot(count(meta, country), 
       aes(x="", y=n, fill=country)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_fill_npg()+
  theme_void()
```

```{r}
b <- ggplot(count(meta, seq_region), 
       aes(x="", y=n, fill=seq_region)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_fill_npg()+
  theme_void()
```

```{r}
c <- ggplot(count(meta, stage), 
       aes(x="", y=n, fill=stage)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_fill_npg()+
  theme_void()
```

```{r}
d <- ggplot(count(meta, platform), 
       aes(x="", y=n, fill=platform)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  scale_fill_npg()+
  theme_void()
```

```{r}
ggarrange(a, b, c, d, ncol = 4, nrow = 1)
#ggsave("output/metapie.pdf", width = 9, height = 5)
```



# Barplot shows the data distribution between timepoints

## All samples
```{r}

meta$age_day <- factor(meta$age_day)
meta$paper <- str_replace(meta$paper, '[ab]$', '')
ggplot(meta) +
  geom_bar(aes(x=age_day, fill=paper))+
  scale_fill_simpsons()+
  theme_minimal()+
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
#ggsave("output/bar_plot.pdf", width = 9, height = 3)
```

## Only control
```{r}
ggplot(filter(meta, treatment == "control")) +
  geom_bar(aes(x=age_day, fill=paper))+
  scale_fill_simpsons()+
  theme_minimal()+
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
#ggsave("output/bar_plot_con.pdf", width = 9, height = 3)
```

## "Rarefy" the metadata
Based on the bar plots shown above, we decide to "rarefy" our metadata on days to avoid the bias introduced by the non-uniform samples size.

```{r}
# days_big <- filter(meta, treatment == "control") %>%
#   count(age_day) %>% filter(n>50) 
# 
# idx_out <- NULL
# 
# for (i in days_big$age_day) {
#   today <- filter(meta, treatment == "control",age_day == i)
#   row.names(today) <- today$sample.id
#   rowname  <- row.names(today)
#   sam.rowname=sample(rowname,50,replace=FALSE)
#   sam.out <- setdiff(rowname, sam.rowname)
#   idx_out <- c(idx_out, sam.out)
# }
# 
# meta_con_rare <- filter(meta, 
#                    treatment == "control",
#                    !(sample.id %in% idx_out))

#write.table(meta_con_rare, "input/metadata.con.rare.txt", quote = F, sep = "\t")

meta_con_rare <- read.table("input/metadata.con.rare.txt",sep = "\t")
meta_con_rare$paper <- str_replace(meta_con_rare$paper, '[ab]$', '')
meta_con_rare$age_day <- factor(meta_con_rare$age_day)

ggplot(meta_con_rare) +
  geom_bar(aes(x=age_day, fill=paper))+
  scale_fill_simpsons()+
  theme_minimal()+
  theme(legend.position = "none")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = .5))
#ggsave("output/bar_plot_con_rare.pdf", width = 9, height = 3)
```




